Supplementary material

### 2021-12-08 submission/manuscript.tex
new version

\begin{abstract}
data using diverse assumptions for a baseline prognostic index of risk
(PI) and the shape of its intereraction with treatment (none, linear or
quadratic). In constant relative treatment effect; models including an
interaction of treatment with the PI; stratification in quarters of the
PI; nonlinear transformations of the PI (restricted cubic splines with
3, 4 and 5 knots); an adaptive approach using Akaike's Information
Criterion. We evaluated predictive performance using root mean squared
error and measures of discrimination and calibration for benefit.
Starting from a base case scenario (sample size 4,250, treatment odds
ratio 0.8, AUC of the PI 0.75), we varied the sample size, the treatment
effect PI by treatment interaction performed better under most
simulation settings. Flexible models required larger sample sizes and
higher AUC of the PI to outperform the linear interaction model. The
adaptive approach tended to select simpler models with smaller sample
sizes and/or worse prediction performance. the PI with treatment is the
optimal risk-based approach to predicting absolute treatment benefit.
\textbf{Objective:} To compare different risk-based methods predicting
individualized treatment effects in RCT simulations. \textbf{Study
Design and Setting:} We simulated each sample we predicted absolute
benefit using: models with the PI and a strength, and the PI's
discriminative ability. \textbf{Results:} Models including a
\textbf{Conclusion:} Under most circumstances, a model with a linear
interaction of
\end{abstract}

@@

For each patient we generated 8 baseline covariates here
\(x_1,\dots,x_4\sim N(0, 1)\) and \(x_5,\dots,x_8\sim B(1, 0.2)\).
Treatment was allocated using a 50:50 split. Outcomes for patients in
the control arm were generated from a logistic regression model
including all baseline covariates. In the base scenarios coefficient
values were such, that the AUC of the logistic regression model was
\(0.75\) and the event rate in the control arm was \(20\%\). Binary
outcomes in the control arm were generated from Bernoulli variables with
true probabilities
\(P(y=1|X, t_x = 0) = \text{expit}(PI)=\frac{e^{PI}}{1+e^{PI}}\).

@@

Second, we considered a model which assumes \emph{constant relative
treatment effect} (constant OR). Hence, absolute benefit is predicted
from \(\hat{\tau}(\bm{x}) = \text{expit}(PI +\log(\text{OR}))\).

@@

Fourth, we used \emph{restricted cubic splines} (RCS) to relax the
linearity assumption on the effect of the linear predictor {[}7{]}. We
considered splines with 3, 4 and 5 knots to compare models with
different levels of flexibility.

@@

\[\text{RMSE}=\frac{1}{n}\sum_{i=1}^n\big(\tau(\bm{x}_i) - \hat{\tau}(\bm{x}_i)\big)^2\]
We compared the discriminative ability of the methods under study using
c-for-benefit {[}8{]}. The c-for-benefit represents the probability that
from two randomly chosen matched patient pairs with unequal observed
benefit, the pair with greater observed benefit also has a higher
predicted benefit. To be able to calculate observed benefit, patients in
each treatment arm are ranked based on their predicted benefit and then
matched 1:1 across treatment arms. \emph{Observed} treatment benefit is
defined as the difference of observed outcomes between the untreated and
the treated patient of each matched patient pair. \emph{Predicted}
benefit is defined as the average of predicted benefit within each
matched patient pair.

@@

The model including a constant relative treatment effect had the lowest
median RMSE in scenarios with a true constant relative treatment effect
(OR = 0.8, N = 4,250 and AUC = 0.75) or with moderate relative
deviations (Figure \ref{fig:rmsebase}; Panel A). However, when we
considered strong linear and quadratic deviations from a constant
relative treatment effect the linear interaction model performed best
(Figure \ref{fig:rmsebase}; Panels B and C). Only in the case of strong
quadratic deviations models including RCS (3 knots) performed equally
well to the linear interaction method. Increasing the number of knots in
RCS resulted in higher RMSE across all scenarios. The adaptive approach
selected a linear interaction model 35\% of the time in the case of true
strong linear deviations, while the majority of the time (56\%) it
selected a simpler constant treatment effect model (Supplementary Table
S7). Similarly, in the case of true strong quadratic deviations, the
adaptive approach selected the RCS (3 knots) model 29\% of the time,
while the simpler linear interaction model was selected 68\% of the
time.

\begin{figure}
\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/rmsebase-1} \caption{RMSE of the considered methods across 500 replications calculated in a simulated superpopulation of size 500,000. The scenario with true constant relative treatment effect had a true prediction AUC of 0.75 and sample size of 4250. Results are presented under moderate linear, strong linear, and strong quadratic deviations from constant relative treatment effects.}\label{fig:rmsebase}
\end{figure}

When we increased the sample size (N = 17,000), the model including a
constant relative treatment effect had the lowest RMSE under the
assumption of true constant relative treatment effects (Figure
\ref{fig:rmsesamplesize}; Panel A). However, when introducing moderate
and strong linear deviations the linear interaction model outperformed
the constant treatment effect model (Figure \ref{fig:rmsesamplesize};
Panels B and C). Furthermore, the linear interaction model was
outperformed by the more flexible RCS models (3 knots) in the case of
strong quadratic deviations. Again, the increased flexibility of RCS
smoothing with higher number of knots resulted in overfitting and worse
predictive accuracy (Figure \ref{fig:rmsesamplesize}; Panel D).

@@

When we increased the AUC of the true prediction model to 0.85, models
including RCS smoothing had the lowest RMSE in the presence of strong
quadratic deviations from the base case of true constant relative
treatment effects (Figure \ref{fig:rmseauc}; Panel D). However, with
milder deviations, the linear interaction model had the lowest RMSE with
the RCS smoothing methods (3 knots) being a close second (Figure
\ref{fig:rmseauc}; Panels B and C). Increasing the number of knots of
RCS smoothing resulted in increased RMSE, which was less pronounced in
the case of strong quadratic deviations. With increasing prediction AUC
we saw an increased tendency of the adaptive approach to select more
complex models compared to the base case scenario. In the case of strong
linear deviations, the linear interaction model was selected the
majority of the time (63\%), while the simpler constant treatment effect
model was selected 26\% of the time (Supplementart Table S7). However,
the increased RMSE of the constant treatment effect model inflated the
the RMSE of the linear interaction model. In the case of strong
quadratic deviations the RCS (3 knots) model was selected 84\% of the
time.

@@

The constant effects model, the linear interaction model and models with
RCS smoothing (3 knots) had the highest median c-for-benefit in the base
case scenario and the scenarios where linear and quadratic deviations
were considered. The constant treatment effect model and the linear
interaction model tended to present much lower variability compared to
all other approaches (Figure \ref{fig:discrimination}). With increasing
number of RCS knots, we observed decreasing median values and increasing
variability of the c-for-benefit in all scenarios.

@@

As for calibration, the linear interaction model had the lowest median
ICI-for-benefit in the majority of scenarios except for the scenarios
where no or moderate linear deviations from the base case were
considered. In that case constant treatment effect models demonstrated
optimal calibration, very comparable to the linear interaction model's
performance, nonetheless (Figure \ref{fig:calibration}). With strong
linear or quadratic deviations, the constant treatment effect model was
poorly calibrated (Figure \ref{fig:calibration}; Panels C and D).

@@

\begin{figure}
\includegraphics[width=1\linewidth]{/home/arekkas/Documents/Projects/arekkas_HteSimulation_XXXX_2021/figures/gusto} \caption{Individualized absolute benefit predictions based on baseline risk when using a constantn treatment effect approach, a linear interaction approach and RCS smoothing using 3,4 and 5 knots. Risk stratified estimates of absolute benefit are presented within quartiles of baseline risk as reference.}\label{fig:gusto}
\end{figure}

@@

The adaptive approach had adequate performance, following closely the
performance of the ``true'' model in most scenarios. However, with
smaller sample sizes it tended to ``miss'' the treatment-risk
interactions and selected simpler models (Supplementary Table S7). This
resulted in increased RMSE variability in these scenarios, especially in
the case of true strong linear deviations from the base case scenario.
This tendency was mitigated when the discriminative ability of the
prediction model was higher. Therefore, in the case of smaller sample
sizes and/or poorly discriminating prediction models, the simpler linear
interaction model is a safer choice for predicting absolute benefits.

Risk-based approaches to predictive HTE estimate treatment benefit as a
function of baseline risk. A limitation of our study is that we assumed
treatment benefit to be a function of baseline risk in the majority of
the simulation scenarios. Nevertheless, our main conclusions did not
change when we generated individual treatment-covariate interactions
(supplementary table/figure x). Future simulation studies could explore
the effect of more extensive deviations from risk-based treatment
effects.

>>>>

\begin{abstract}
\textbf{Objective:} To compare different risk-based methods predicting
individualized treatment effects in RCTs. \textbf{Study Design and
Setting:} We simulated data using diverse assumptions for a baseline
prognostic index of risk (PI), the shape of its interaction with
treatment (none, linear, quadratic or non-monotonic) and the magnitude
of treatment-related harms. In each sample we predicted absolute benefit
using: models with a constant relative treatment effect; stratification
in quarters of the PI; models including a linear interaction of
treatment with the PI; models including an interaction of treatment with
a restricted cubic spline (RCS) transformation of the PI; an adaptive
approach using Akaike's Information Criterion. We evaluated predictive
performance using root mean squared error and measures of discrimination
and calibration for benefit. Starting from a base scenario (sample size
4,250, treatment odds ratio 0.8, AUC of the PI 0.75), we varied the
sample size, the treatment effect size, and the PI's AUC.
\textbf{Results:} Models including a PI by treatment interaction
performed best with smaller sample sizes and/or lower AUC of the PI.
Otherwise, RCS (3 knots) models proved more robust for the majority of
the deviation settings. The adaptive approach was unstable with smaller
sample sizes. \textbf{Conclusion:} Depending on the setting, the linear
interaction or the RCS (3 knots) model should be preferred.
\end{abstract}

@@

For each patient we generated 8 baseline covariates
\(x_1,\dots,x_4\sim N(0, 1)\) and \(x_5,\dots,x_8\sim B(1, 0.2)\).
Treatment was allocated using a 50:50 split. Outcomes for patients in
the control arm were generated from a logistic regression model
including all baseline covariates. In the base scenarios coefficient
values were such, that the AUC of the logistic regression model was
\(0.75\) and the event rate in the control arm was \(20\%\). Binary
outcomes in the control arm were generated from Bernoulli variables with
true probabilities
\(P(y=1|X, t_x = 0) = \text{expit}(PI)=\frac{e^{PI}}{1+e^{PI}}\).

@@

Second, we considered a model which assumes \emph{constant relative
treatment effect} (constant odds ratio). Hence, absolute benefit is
predicted from
\(\hat{\tau}(\bm{x}) = \text{expit}(PI +\log(\text{OR}))\).

@@

Fourth, we used \emph{restricted cubic splines} (RCS) to relax the
linearity assumption on the effect of the linear predictor {[}7{]}. We
considered splines with 3 (RCS-3), 4 (RCS-4) and 5 (RCS-5) knots to
compare models with different levels of flexibility.

@@

\[\text{RMSE}=\sqrt{\frac{1}{n}\sum_{i=1}^n\big(\tau(\bm{x}_i) - \hat{\tau}(\bm{x}_i)\big)^2}\]
We compared the discriminative ability of the methods under study using
c-for-benefit {[}8{]}. The c-for-benefit represents the probability that
from two randomly chosen matched patient pairs with unequal observed
benefit, the pair with greater observed benefit also has a higher
predicted benefit. To be able to calculate observed benefit, patients in
each treatment arm are ranked based on their predicted benefit and then
matched 1:1 across treatment arms. \emph{Observed} treatment benefit is
defined as the difference of observed outcomes between the untreated and
the treated patient of each matched patient pair. \emph{Predicted}
benefit is defined as the average of predicted benefit within each
matched patient pair.

@@

The linear interaction model outperformed all RCS methods in terms of
RMSE in scenarios with true constant relative treatment effect (OR =
0.8, N = 4,250 and AUC = 0.75), strong linear and even strong quadratic
deviations (Figure \ref{fig:rmsebase}). However, with non-monotonic
deviations errors of the linear interaction model increased
substantially, especially in the presence of treatment-related harms. In
these scenarios, RCS-3 proved to be quite robust outperforming all other
methods. The constant treatment effect approach had overall best
performance under true constant treatment effect settings, but was
sensitive to all considered deviations, resulting in increased error
rates. Finally, the adaptive approach had comparable performance to the
best-performing method in each scenario. However, we observed increased
error variability in the case of linear and non-monotonic deviations,
especially for moderate or strong treatment-related harms.

In the case of true constant relative treatment effects, the adaptive
approach selected the constnant effect model the majority of the time,
even in the presence of treatment-related harms (96\% of the time with
no harms to 88\% with strong harms). With strong linear or quadratic
deviations and increasing treatment-related harms the adaptive approach
increasingly favored the linear interaction model. In the case of true
non-monotonic deviations, we observed increasing selection frequencies
for RCS-3 with increasing treatment harms (from 32\% with no harms to
53\% with strong harms), whereas selection frequencies for the linear
interaction model dropped from 40\% with no harms to 7\% with strong
harms. Finally, in the case of strong linear and non-monotonic
deviations selection frequencies of the constant treatment effect model
remained high despite increasing treatment harms (Supplement, Table XX).

\begin{figure}
\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/rmsebase-1} \caption{RMSE of the considered methods across 500 replications calculated in a simulated super-population of size 500,000. The scenario with true constant relative treatment effect had a true prediction AUC of 0.75 and sample size of 4250. Results are presented under moderate linear, strong linear, and strong quadratic deviations from constant relative treatment effects.}\label{fig:rmsebase}
\end{figure}

Increasing the sample size to 17,000 favored RCS-3 the most, as it
achieved lowest or close to lowest RMSE across all scenarios (Figure
\ref{fig:rmsesamplesize}). Especially in cases of strong quadratic and
non-monotonic deviations RCS-3 had lower error rates (median RMSE 0.011
for strong quadratic deviations and 0.010 for non-monotonic deviations
with no treatment-related harms) compared to the linear interaction
approach (median RMSE 0.013 and 0.014, respectively), regardless of
treatment-related harms strength. The issues with the large error
variability of the adaptive approach improved with larger sample sizes.

The adaptive approach tended to increasingly favor smoother methods
(especially RCS-3) with increasing treatment-related harms (see
Supplement, Table XX). However, in the case of true strong quadratic
deviations the opposite was observed: selection frequency of the linear
interaction model increased from 31\% (no harms) to 50\% (strong harms)
whereas for RCS-3 decreased from 52\% (no harms) to 34\% (strong harms).

@@

When we increased the AUC of the true prediction model to 0.85 (OR = 0.8
and N = 4,250) the constant effect model outperformed all other methods
in the case of true constant treatment effects, but proved to be the
least robust to deviations. Again, RCS-3 had the lowest error rates in
the case of strong quadratic or monotonic deviations and very comparable
performance to the best-performing linear interaction model in the case
of strong linear deviations (median RMSE 0.016 for RCS-3 compared to
0.014 for the linear interaction model). The adaptive approach, though
it performed similar to the best performing method in each scenario, on
average, had increased variability in error rates in the case of strong
linear and non-monotonic deviations. In these scenarios, the adaptive
approach often selected the constant treatment effect model (23\% and
25\% in the strong linear and non-monotonic deviation scenarios without
treatment-related harms, respectively).

@@

In terms of discrimination for benefit (OR = 0.8, N = 4,250 and AUC =
0.75), all methods performed similarly, on average (Figure
\ref{fig:discrimination}). In the case of non-monotonic deviations, the
constant effect model had much lower discriminative performance compared
to the rest of the methods (median AUC of 0.04 for the constant effects
model compared to the best-performing RCS-3 with 0.02). The linear
interaction model was the most stable compared to the other methods in
terms of error variability. With increasing number of RCS knots, we
observed decreasing median values and increasing variability of the
c-for-benefit in all scenarios.

@@

In terms of calibration for benefit, the constant effects model
outperformed all other models in the case of true constant treatment
effects, but was miscalibrated for all deviation scenarios (Figure
\ref{fig:calibration}). The linear interaction model showed best or
close to best calibration across all scenarios and only showed worse
calibration compared to RCS-3 in the case of non-monotonic deviations
and treatment-related harms. The adaptive approach was worse calibrated
in scenarios with strong linear and non-monotonic deviations compared to
the linear interaction model and RCS-3.

The results from all individual scenarios can be explored online at
\url{https://arekkas.shinyapps.io/simulation_viewer/}.

@@

\begin{figure}
\includegraphics[width=1\linewidth]{/home/arekkas/Documents/Projects/arekkas_HteSimulation_XXXX_2021/figures/gusto} \caption{Individualized absolute benefit predictions based on baseline risk when using a constant treatment effect approach, a linear interaction approach and RCS smoothing using 3,4 and 5 knots. Risk stratified estimates of absolute benefit are presented within quartiles of baseline risk as reference.}\label{fig:gusto}
\end{figure}

@@

The adaptive approach had adequate performance, following closely on
average the performance of the ``true'' model in most scenarios.
However, with smaller sample sizes it tended to ``miss'' the
treatment-risk interactions and selected simpler models (Supplementary
Table S7). This resulted in increased RMSE variability in these
scenarios, especially in the case of true strong linear or non-monotonic
deviations from the base case scenario. Therefore, in the case of
smaller sample sizes the simpler linear interaction model is a safer
choice for predicting absolute benefits.

Risk-based approaches to predictive HTE estimate treatment benefit as a
function of baseline risk. A limitation of our study is that we assumed
treatment benefit to be a function of baseline risk in the majority of
the simulation scenarios. We attempted to address that by introducing
constant moderate and strong treatment-related harms, applied on the
absolute scale. Also, we considered a small number of scenarios with
true treatment-covariate interactions, in which our main conclusions
remained the same (Supplement, XX). Future simulation studies could
explore the effect of more extensive deviations from risk-based
treatment effects.

### 2021-12-08 submission/supplement.rmd
new version

```{r tab1, echo=FALSE, eval=TRUE, warning=FALSE}
 asd <- read_csv(
  here("data/processed/analysisIds.csv"),
  col_types = "iffiiddddddddddddd"
) %>%
  select(-"type")

kableExtra::kable(
  asd,
  format = "latex",
  escape = FALSE,
  longtable   = TRUE,
  align       = "r",
  booktabs    = TRUE,
  caption     = "",
  col.names   = c(
    "Scenario",
    "Effect",
    "N",
    "AUC",
    paste0("b", 0:8),
    paste0("g", 0:2),
    "c"
  ),
  digits      = 2,
  format.args = list(
    big.mar      = ",",
    decimal.mark = "."
  )
) %>%
  kableExtra::group_rows(
    group_label = "Constant treatment effect",
    start_row   = 10,
    italic      = TRUE,
    end_row     = 27
  ) %>%
  kableExtra::group_rows(
    group_label = "Linear deviation",
    start_row   = 28,
    italic      = TRUE,
    end_row     = 45
  ) %>%
  kableExtra::group_rows(
    group_label = "Quadratic deviation",
    start_row   = 46,
    italic      = TRUE,
    end_row     = 63
  ) %>%
  kableExtra::landscape() %>%
  kableExtra::add_header_above(
    c(
      "Analysis ID"           = 4,
      "Baseline risk"         = 9,
      "True treatment effect" = 4
      )
  )
```

\newpage
```{r tab2, echo=FALSE, eval=TRUE, warning=FALSE}
asd <- read_csv(
  here("data/processed/analysisIdsInteractions.csv"),
  col_types = "iffiidddddddddddddd"
) %>%
  select(-c("type", "g0"))

  knitr::kable(
    asd,
    format      = "latex",
    longtable   = TRUE,
    align       = "r",
    booktabs    = TRUE,
    col.names   = c(
      "Scenario",
      "Effect",
      "N",
      "AUC",
      paste0("b", 0:8),
      "g1", "g2", "g5", "g6"
    ),
    digits      = 2,
    format.args = list(
      big.mar      = ",",
      decimal.mark = "."
    ),
    caption    = ""
  ) %>%
  kableExtra::landscape() %>%
  kableExtra::add_header_above(
    c(
      "Analysis ID"           = 4,
      "Baseline risk"         = 9,
      "Coefficient in\ntreatment arm" = 4
      )
  )
```

# Results in scenarios with interactions

When we considered a set of 4 true linear treatment-covariate interactions the
model containing a linear interaction with the prognostic index had the lowest
median RMSE. We observed an increasing trend in prediction errors with
increasing interaction intensity (Figure \@ref(fig:rmseInteractions)). The
model with restricted cubic spline smoothing (3 knots) had very comparable
performance to the linear interaction model. Increasing the flexibility of the
smooth methods resulted in increasing median RMSE. These results may be
explained by the fact that the interactions considered were linear, thus
favoring the linear interaction model. More flexible approaches may be better
suited for higher-order treatment-covariate interactions. Finally, the adaptive
approach had adequate performance under all scenarios, resembling the
performance of the best-performing approach every time.

```{r rmseInteractions, echo=FALSE, fig.cap="Linear and quadratic deviations from the base-case scenario of constant relative effect (OR=0.8)", fig.show="hold", out.width = '100%'}
knitr::include_graphics(here::here("figures/rmse_interactions.png"))
```

The constant treatment effects model, the linear interaction model and the model
with RCS smoothing (3 knots) had the highest c-for-benefit across all scenarios
(Figrue \@ref(fig:discriminationInteractions)). RCS smoothing with 4 or 5 knots
did not improve performance. On the contrary, we observed an increasing trend in
c-for-benefit variability, as was the case in the main text. The adaptive
apprach again had statisfactory performance.

```{r discriminationInteractions, echo=FALSE, fig.cap="Linear and quadratic deviations from the base-case scenario of constant relative effect (OR=0.8)", fig.show="hold", out.width = '100%'}
knitr::include_graphics(here::here("figures/discrimination_interactions.png"))
```

Despite the very similar performance in terms of prediction errors, the linear
interaction model resulted in better-calibrated benefit predictions compared to
the rest of the methods (Figure \@ref(fig:calibrationInteractions)). The constant
treatment effects model had the highest median ICI-for-benefit across all
scenario settings, which became more pornounced with increasing
treatment-covariate interaction intensity.

```{r calibrationInteractions, echo=FALSE, fig.cap="Linear and quadratic deviations from the base-case scenario of constant relative effect (OR=0.8)", fig.show="hold", out.width = '100%'}
knitr::include_graphics(here::here("figures/calibration_interactions.png"))
```

# Case study

We fitted a logistic regression model to predict 30-day mortality after a
myocardial infarction using data from the GUSTO-I trial. The risk factors
modeled were age, Killip class, systolic blood pressure heart rate, primary MI
and MI location. We restricted systolic blood pressure values to 120 mmHg and
modeled heart rate using linear spline transformation with a knot at 50 beats
per minute. A brief overview of the distribution of baseline characteristics is
presented in Table \@ref(tab:table1). The coefficients of the model can be seen
in Table \@ref(tab:modelCoefficients). We used this model in the manuscript to
derive baseline risk predictions to all patients of the study. For all patients
we set the treatment indicator to 0, assuming a constant relative treatment
effect. The model displayed very good internal validity (Table
\@ref(tab:internalValidity)).

```{r table1, echo=FALSE, warning=FALSE, message=FALSE}
load(here::here("data/raw/gusto.rda"))
gusto <- gusto %>%
  tibble() %>%
  filter(!is.na(tpa))

gustoTable1 <- gusto %>%
  mutate(
    tpa = factor(
      tpa,
      levels = c(0, 1),
      labels = c(
        "Tissue plasminogen activator",
        "Streptokinase"
      )
    ),
    Age = age,
    Outcome = factor(
      day30,
      levels = c(0, 1),
      labels = c(
        "Alive",
        "Dead"
      )
    )
  )

res <- table1::table1(
  ~ sex + Age + Killip + sysbp + + miloc + pulse + pmi + Outcome | tpa,
  data = gustoTable1,
  caption = "Baseline characteristics of the patients included in GUSTO-I trial."
)

table1::t1kable(res)
```

```{r modelCoefficients, echo=FALSE, warning=FALSE, message=FALSE}
gustoModelData <- gusto %>%
  rename(
    "outcome" = "day30",
    "treatment" = "tpa"
  )

prediction <- glm(
  outcome ~ treatment + age + Killip + pmin(sysbp, 120) + rms::lsp(pulse, 50) + pmi + miloc,
  data = gustoModelData,
  maxit = 99,
  family = "binomial"
)

predictionSummary <- summary(prediction)
table <- predictionSummary$coefficients

rownames(table)[2] <- "Treatment"
rownames(table)[3] <- "Age"
rownames(table)[4] <- "KILLIP (II)"
rownames(table)[5] <- "KILLIP (III)"
rownames(table)[6] <- "KILLIP (IV)"
rownames(table)[7] <- "Systolic blood pressure"
rownames(table)[8] <- "Pulse rate"
rownames(table)[9] <- "Pulse rate'"
rownames(table)[10] <- "Prior MI (yes)"
rownames(table)[11] <- "MI location (other)"
rownames(table)[12] <- "MI location (anterior)"

 knitr::kable(
    table,
    format      = "latex",
    align       = "r",
    booktabs    = TRUE,
    digits      = 4,
    format.args = list(
      big.mar      = ",",
      decimal.mark = "."
    ),
    caption    = "Coefficients of the prediction model for 30-day mortality after myocardial infarction (MI)."
  )

```

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Smooth calibration of the derived prediction model."}
predictionStats <- rms::val.prob(
  y = gustoModelData$outcome, 
  logit = prediction$linear.predictors,
  logistic.cal = FALSE,
  statloc = FALSE
)
```

```{r internalValidity, echo=FALSE, warning=FALSE, message=FALSE}
table <- tibble(
  Metric = c(
    "AUC",
    "Brier score",
    "Emax",
    "E90",
    "Eavg"
  ),
  Value = c(
    predictionStats[2],
    predictionStats["Brier"],
    predictionStats["Emax"],
    predictionStats["E90"],
    predictionStats["Eavg"]
  )
)

 knitr::kable(
    table,
    format      = "latex",
    align       = "l",
    booktabs    = TRUE,
    digits      = 4,
    format.args = list(
      big.mar      = ",",
      decimal.mark = "."
    ),
    caption    = "Evaluation statistics of the prediction model for 30-day mortality in GUSTO-I."
  ) 
```

We used the derived prediction model to predict benefit of tPA to 30-day
mortality using the data at hand. The models are presented in the main
manuscript. All approaches displayed comaprable discrimination for benefit, with
the constant treatment effect model, the linear interaction model and the RCS (5
knots) model displaying the highest performance (Table
\@ref(tab:gustoPerformanceMetrics)). However, the linear interaction model and
the RCS (5 knots) model were worse calibrated. The constant treatment effect
model also had the best performance among the considered methods in terms of
calibration for benefit.

```{r gustoPerformanceMetrics, echo=FALSE, warning=FALSE, message=FALSE}
metrics <- readr::read_csv(
  here::here("data/processed/gustoPerformanceMetrics.csv")
)

 knitr::kable(
    metrics,
    format      = "latex",
    align       = "l",
    booktabs    = TRUE,
    digits      = 4,
    format.args = list(
      big.mar      = ",",
      decimal.mark = "."
    ),
    col.names = c(
      "Model",
      "C",
      "ICI",
      "AIC"
    ),
    caption    = "Performance metrics of the different methods used to predict benefit of tPA on 30-day mortality using data from the GUSTO-I trial."
  ) 
```

>>>>

```{r tab1, echo=FALSE, eval=TRUE, warning=FALSE, message=FALSE}
 asd <- read_csv(
  here("data/processed/analysisIds.csv")
) %>%
  select(-"type")
kableExtra::kable(
  asd,
  format = "latex",
  escape = FALSE,
  longtable   = TRUE,
  align       = "r",
  booktabs    = TRUE,
  caption     = "",
  col.names   = c(
    "Scenario",
    "Base",
    "N",
    "AUC",
    "Treatment-related harm",
    paste0("b", 0:8),
    paste0("g", 0:2),
    "c",
    "Average benefit"
  ),
  digits      = 2,
  format.args = list(
    big.mar      = ",",
    decimal.mark = "."
  )
) %>%
  kableExtra::landscape() %>%
  kableExtra::add_header_above(
    c(
      "Analysis ID"           = 5,
      "Baseline risk"         = 9,
      "True treatment effect" = 4
      )
  ) %>%
  kableExtra::kable_styling(font_size = 7)
```
